Aligning species range data to better serve science and conservation

### 2015-12-28 data_explore/data_explore.Rmd
updates to manuscript; also data_explore still needs to be updated with new species lists and the like.

spp_list <- spp_list %>%
  filter(!is.na(sm_perc)) %>%
  ### will ditch the ones where no IUCN polygon exists - or are these species
  ### whose ranges only occur in cells beyond the LOICZID grid?
  filter(!str_detect(spatial_source, 'subpop') & !str_detect(spatial_source, 'alias'))
spp_list <- spp_list %>%
  left_join(read.csv(file.path(dir_data, 'sciname_lookups/spp_groupnames.csv'), 
                     stringsAsFactors = FALSE),
            by = 'spp_group')

>>>>

spp_list <- spp_list %>%
  filter(!is.na(sm_perc)) %>%
  ### will ditch the ones where no IUCN polygon exists - or are these species
  ### whose ranges only occur in cells beyond the LOICZID grid?
  filter(!str_detect(spatial_source, 'subpop') & !str_detect(spatial_source, 'alias'))
spp_list <- spp_list %>%
  left_join(read.csv(file.path(dir_data, 'sciname_lookups/spp_groupnames.csv'), 
                     stringsAsFactors = FALSE),
            by = 'spp_group') %>%
  mutate(spp_group = ifelse(str_detect(spp_group, 'corals'), 'corals', spp_group))
    ### lump all corals together just for ease of handling

### 2015-12-28 ms_4_figs.Rmd
updates to manuscript; also data_explore still needs to be updated with new species lists and the like.

### Figure 1 (a, b):

@@

![](figures/bar_by_taxa_assessed.png)

Number and proportion of species, listed by taxa, included in each dataset: IUCN, AquaMaps, or both. (a) All species with distribution maps; (b) IUCN Red List-assessed species with distribution maps.

### Figure 2 (a, b):

@@

Distribution alignment vs Extent alignment for `r nrow(spp_map_pairs)` species mapped in both Aquamaps and IUCN datasets.

@@

* The upper right quadrant (quadrant 1) comprises species whose maps largely agree (better than median value) in both spatial distribution and the area of described ranges.
(n = `r q1_count`; `r round(q1_count/nrow(spp_map_pairs) * 100, 1)` %)
* The upper left quadrant (quadrant 2) comprises species whose maps agree well in distribution, but disagree in extent.
(n = `r q2_count`; `r round(q2_count/nrow(spp_map_pairs) * 100, 1)` %)
* The lower right quadrant (quadrant 3) includes species for which the paired maps generally agree in range extent, but disagree on where those ranges lie - a more problematic mismatch than that indicated by quadrant 2.
(n = `r q3_count`; `r round(q3_count/nrow(spp_map_pairs) * 100, 1)` %)
* The lower left quadrant (quadrant 4) indicates species for which the map pairs fail to agree in both area and distribution. 
(n = `r q4_count`; `r round(q4_count/nrow(spp_map_pairs) * 100, 1)` %)

### Figure 3 (b, c):

@@

![](figures/barchart_spp_risk_quads.png)

Spatial alignment of paired-map species by (a) taxonomic group and (b) extinction risk category.

### Figure 5 (a, b):

![](figures/AMThreshold.png)

![](figures/ribbon_AMThreshold.png)

AquaMaps distribution map extent remaining after applying a presence threshold.  (a) A 40% threshold applied to all species in the AquaMaps dataset shows a mean loss of XXX, with a wide distribution in which some species lose nearly all of their apparent range.  (b) Mean (median) remaining extent at increments of presence threshold. Dark grey ribbon includes 25% to 75% quantiles, while light grey ribbon includes 9% to 91% quantiles.

### Figure 6

>>>>

### Figure 1(a)?:

@@

Number and proportion of species, listed by taxa, included in each dataset: IUCN, AquaMaps, or both.

### Figure 1 (b, c):

@@

For each paired-map species, we calculated two dimensions of spatial alignment: _distribution alignment_, which we defined as the proportion of the smaller range intersecting the larger range (where on the map); and _extent alignment_, which we defined as the ratio of the smaller range to the larger range (how much of the map).  For a species whose distribution is well understood and described in both datasets, we would expect to see a value near 100% for each dimension of alignment.

@@

* The upper right quadrant (quadrant 1) comprises species whose maps largely agree (better than median value) in both spatial distribution and the extent of described ranges. 
(n = `r q1_count`; `r round(q1_count/nrow(spp_map_pairs) * 100, 1)` %)
* The upper left quadrant (quadrant 2) comprises species whose maps agree well in distribution, but disagree in extent.
(n = `r q2_count`; `r round(q2_count/nrow(spp_map_pairs) * 100, 1)` %)
* The lower right quadrant (quadrant 3) includes species for which the paired maps generally agree in range extent, but disagree on where those ranges lie - a more worrisome mismatch than that indicated by quadrant 2.
(n = `r q3_count`; `r round(q3_count/nrow(spp_map_pairs) * 100, 1)` %)
* The lower left quadrant (quadrant 4) indicates species for which the map pairs agree poorly in both area and distribution. 
(n = `r q4_count`; `r round(q4_count/nrow(spp_map_pairs) * 100, 1)` %)

### Figure 3 (b):

@@

Spatial alignment of paired-map species by taxonomic group.

### Figure 4
